\section{Methods}

Full atom simulations of DNA molecules exist but are computationally very heavy -- some methods even simulate the individual atoms of the solvent. These models yield of course a very high degree of detail, but the low computational efficiency forces these models to be useful only in describing small DNA oligomers (20-30 bp). LAMMPS \cite{plimpton2007lammps}, a free and open source molecular simulation framework, is perhaps the most widespread of these tools, but other examples are the force-field based atomistic models like CHARMM \cite{brooks2009charmm} and AMBER \cite{cheatham1999modified}.

In 2007, the group of De Pablo \cite{knotts2007coarse} applied the technique of coarse graining to a molecular simulation of DNA: instead of simulating all atoms, a DNA monomer is represented by three sites or `beads': the sugar, the phosphate and the base. For simulations where one is mainly interested in the timescale behaviour or global dynamics of the DNA molecule this 3SPN (three-sites-per-nucleotide) model is a very good approximation: it is computationally cheap (and easy to optimize) while describing the DNA helical structure quite good. 

We implemented the 3SPN model with a few changes of our own: we slightly modified the way base pairing forces were calculated and we custom-fitted some constants to better describe the zipping and unzipping dynamics of ssDNA hairpins. We also used the technique of spatial partitioning to optimize the algorithmic complexity.

Our code base is publicly available and is hosted on GitHub: \href{https://github.com/RoaldFre/DNA}{github.com/RoaldFre/DNA} for the foreseeable future. The model is implemented in C with an optional OpenGL rendering engine.

The interested reader can find the full package containing everything needed to reproduce the material presented in this paper at the above location.


\input{structure}

\input{interactions}

\input{spacePartitioning}

\input{dynamics}


